---
title: "Principal Component Analysis"
author: "Dataset training.ids()"
output:
  html_document: default
---

This dataset contains 30 individuals and 91 variables.

------------------------------------------------------------------------

```{r, echo = FALSE}
library(FactoMineR)
load('E:/sDrive/Git/Eps2Fold/Workspace.RData')
```

### 1. Study of the outliers

The analysis of the graphs does not detect any outlier.

------------------------------------------------------------------------

### 2. Inertia distribution

The inertia of the first dimensions shows if there are strong relationships between variables and suggests the number of dimensions that should be studied.

The first two dimensions of analyse express **81.86%** of the total dataset inertia ; that means that 81.86% of the individuals (or variables) cloud total variability is explained by the plane. This percentage is high and thus the first plane represents an important part of the data variability. This value is strongly greater than the reference value that equals **16.06%**, the variability explained by this plane is thus highly significant (the reference value is the 0.95-quantile of the inertia percentages distribution obtained by simulating 12330 data tables of equivalent size on the basis of a normal distribution).

From these observations, it is probably not useful to interpret the next dimensions.

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
par(mar = c(2.6, 4.1, 1.1, 2.1))
ggplot2::ggplot(cbind.data.frame(x=1:nrow(res$eig),y=res$eig[,2])) + ggplot2::aes(x=x, y=y)+ ggplot2::geom_col(fill="blue") + ggplot2::xlab("Dimension") + ggplot2::ylab("Percentage of variance") + ggplot2::ggtitle("Decomposition of the total inertia") + ggplot2::theme_light() + ggplot2::theme(plot.title = ggplot2::element_text(hjust =0.5)) + ggplot2::scale_x_continuous(breaks=1:nrow(res$eig))
```

**Figure 2 - Decomposition of the total inertia** *The first factor is major: it expresses itself 65.5% of the data variability.* *Note that in such a case, the variability related to the other components might be meaningless, despite of a high percentage.*

An estimation of the right number of axis to interpret suggests to restrict the analysis to the description of the first 3 axis. These axis present an amount of inertia greater than those obtained by the 0.95-quantile of random distributions (93.54% against 22.87%). This observation suggests that only these axis are carrying a real information. As a consequence, the description will stand to these axis.

------------------------------------------------------------------------

### 3. Description of the plane 1:2

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("18", "21", "28", "19", "8", "11", "25", "10", "12", "5", "6", 
"7", "30", "16", "9", "2", "20")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = c(1,2), choix = 'ind', invisible = 'quali', title = '', cex = cex)
```

**Figure 3.1 - Individuals factor map (PCA)** *The labeled individuals are those with the higher contribution to the plane construction.*

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("ids-270", "ids-269", "ids-271", "ids-268", "ids-272", "ids-267", 
"ids-273", "ids-266", "ids-274", "ids-241", "ids-240", "ids-242", 
"ids-275", "ids-265", "ids-239", "ids-243", "ids-276")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = 1:2, choix = 'var', title = '', cex = cex)
```

**Figure 3.2 - Variables factor map (PCA)** *The labeled variables are those the best shown on the plane.*

------------------------------------------------------------------------

The **dimension 1** opposes individuals such as *12*, *5* and *21* (to the right of the graph, characterized by a strongly positive coordinate on the axis) to individuals such as *19*, *10* and *11* (to the left of the graph, characterized by a strongly negative coordinate on the axis).

The group in which the individuals *12* and *5* stand (characterized by a positive coordinate on the axis) is sharing :

-   high values for the variables *ids-270*, *ids-271* and *ids-269* (variables are sorted from the strongest).
-   low values for the variables *ids-296*, *ids-297*, *ids-295* and *ids-298* (variables are sorted from the weakest).

The group in which the individual *21* stands (characterized by a positive coordinate on the axis) is sharing :

-   low values for variables like *ids-303*, *ids-304*, *ids-302*, *ids-305*, *ids-301*, *ids-300*, *ids-306*, *ids-299*, *ids-298* and *ids-307* (variables are sorted from the weakest).

The group in which the individuals *19*, *10* and *11* stand (characterized by a negative coordinate on the axis) is sharing :

-   low values for variables like *ids-228*, *ids-229*, *ids-227*, *ids-230*, *ids-226*, *ids-231*, *ids-225*, *ids-232*, *ids-259* and *ids-260* (variables are sorted from the weakest).

Note that the variables *ids-258*, *ids-259*, *ids-260*, *ids-261*, *ids-262*, *ids-263*, *ids-264*, *ids-265*, *ids-266* and *ids-267* are highly correlated with this dimension (respective correlation of 0.91, 0.93, 0.94, 0.95, 0.95, 0.94, 0.94, 0.94, 0.95, 0.95, 0.96, 0.96, 0.96, 0.95, 0.95, 0.94, 0.94, 0.93, 0.92, 0.91). These variables could therefore summarize themselve the dimension 1.

------------------------------------------------------------------------

The **dimension 2** opposes individuals such as *6*, *2*, *20*, *8* and *7* (to the top of the graph, characterized by a strongly positive coordinate on the axis) to individuals such as *16* and *28* (to the bottom of the graph, characterized by a strongly negative coordinate on the axis).

The group in which the individuals *6*, *20* and *8* stand (characterized by a positive coordinate on the axis) is sharing :

-   low values for the variables *ids-281*, *ids-280*, *ids-282*, *ids-279*, *ids-283*, *ids-278* and *ids-277* (variables are sorted from the weakest).

The group in which the individuals *2* and *7* stand (characterized by a positive coordinate on the axis) is sharing :

-   high values for variables like *ids-250*, *ids-249*, *ids-251*, *ids-248*, *ids-247*, *ids-246*, *ids-252*, *ids-245*, *ids-244* and *ids-236* (variables are sorted from the strongest).

The group in which the individuals *16* and *28* stand (characterized by a negative coordinate on the axis) is sharing :

-   high values for the variables *ids-288*, *ids-289*, *ids-287*, *ids-290*, *ids-286* and *ids-291* (variables are sorted from the strongest).

------------------------------------------------------------------------

### 4. Description of the dimension 3

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("27", "25", "4", "22", "11", "9", "29", "3", "18", "17", "21", 
"26", "28", "13", "12", "7", "10")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = c(3,4), choix = 'ind', invisible = 'quali', title = '', cex = cex)
```

**Figure 4.1 - Individuals factor map (PCA)** *The labeled individuals are those with the higher contribution to the plane construction.*

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("ids-223", "ids-222", "ids-224", "ids-221", "ids-225", "ids-220", 
"ids-290", "ids-226", "ids-289", "ids-291", "ids-292", "ids-227", 
"ids-293", "ids-288", "ids-294", "ids-295", "ids-296")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = 3:4, choix = 'var', title = '', cex = cex)
```

**Figure 4.2 - Variables factor map (PCA)** *The labeled variables are those the best shown on the plane.*

------------------------------------------------------------------------

The **dimension 3** opposes individuals such as *9*, *18* and *29* (to the right of the graph, characterized by a strongly positive coordinate on the axis) to individuals such as *4*, *3*, *25* and *17* (to the left of the graph, characterized by a strongly negative coordinate on the axis).

The group in which the individuals *9*, *18* and *29* stand (characterized by a positive coordinate on the axis) is sharing :

-   high values for variables like *ids-290*, *ids-289*, *ids-291*, *ids-292*, *ids-288*, *ids-293*, *ids-294*, *ids-295*, *ids-220* and *ids-296* (variables are sorted from the strongest).

The group in which the individuals *4*, *3*, *25* and *17* stand (characterized by a negative coordinate on the axis) is sharing :

-   low values for variables like *ids-223*, *ids-222*, *ids-224*, *ids-221*, *ids-220*, *ids-225*, *ids-226*, *ids-227*, *ids-228* and *ids-229* (variables are sorted from the weakest).

------------------------------------------------------------------------

### 5. Classification

```{r, echo = FALSE}
res.hcpc = HCPC(res, nb.clust = -1, graph = FALSE)
```

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("18", "21", "28", "19", "8", "11", "25", "10", "12", "5", "6", 
"7", "30", "16", "9", "2", "20")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.HCPC(res.hcpc, choice = 'map', draw.tree = FALSE, select = drawn, title = '')
```

**Figure 5 - Ascending Hierarchical Classification of the individuals.** *The classification made on individuals reveals 3 clusters.*

The **cluster 1** is made of individuals such as *6*, *8*, *9*, *10*, *11*, *19* and *20*. This group is characterized by :

-   variables whose values do not differ significantly from the mean.

The **cluster 2** is made of individuals such as *16*, *25*, *28* and *30*. This group is characterized by :

-   variables whose values do not differ significantly from the mean.

The **cluster 3** is made of individuals such as *2*, *5*, *7*, *12*, *18* and *21*. This group is characterized by :

-   variables whose values do not differ significantly from the mean.

------------------------------------------------------------------------

## Annexes

```{r, comment = ''}
res.hcpc$desc.var
```

**Figure 6 - List of variables characterizing the clusters of the classification.**